---
title: "Assignment III"
author: "Dana Jensen"
date: "February 17, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# setting working directory and loading data, packages etc.
setwd("C:/Users/Dana/Desktop/METHODS II/CLASS EXERCIZES")
library(psych)
library(ggplot2)
parametres = read.table("parametres.txt", header = FALSE)
fmri<-as.matrix(read.csv("aud_fmri_data37.csv", header=FALSE))
fmri2<-ts(fmri)
fmrides<-as.matrix(read.csv("aud_fmri_design.csv", header=FALSE))
fmrides2<-ts(fmrides)
```

Import the realignment parameters into R.
3.a. Make a lineplot of the realignment parameters in R.
```{r}

# making a dataframe
parametres$Session = c(1:400)
lineplotdata <- stack(parametres, c(V1, V2, V3, V4, V5, V6)) 
lineplotdata$session = c(1:400, 1:400, 1:400, 1:400, 1:400, 1:400)

# lineplot of the realignment parameters
lineplot <- ggplot(lineplotdata, aes(session, values, group = ind, colour = ind)) + geom_line()

#lineplot of realignment parametres
lineplot

```
3.b. How far has the participant moved for each dimension during the experiment (Hint: use "apply()" to
run a function across columns)?
```{r}

# finding the range of how far the participant moved
# 2 indicates rows, a 1 would indicate columns
ranges = apply(parametres, 2, range)
ranges1 = ranges[1,-7] - ranges[2,-7]

# length of movement
print(round(ranges1, 2))

# sum of dimensions 1:3 vs 4:6
sum(ranges1[1:3])
sum(ranges1[4:6])

```
There was more movement in the first three dimensions compared to the last three dimensions, -1.47 vs. -0.02.

3.c. Are any of the realignment parameters significantly correlated with the fMRI model (same model as
used in exercise 2)?
```{r}

# converting to dataframe
fmrides<-data.frame(fmrides)

# testing for correlation
psych::corr.test(parametres,fmrides)

```
V1, the first story, and V5, one of the dimensions of the fMRI, are significantly correlated: r= 0.17, p<.05


Remove linear effects of time from the realignment parameters (hint: fit a line and use residuals).
```{r}

# fitting a linear model to the data to find residuals
model = lm(values ~ session, lineplotdata)
summary(model)

# extracting residuals for each dimension
r1 = model$residuals[lineplotdata$ind == "V1"]
r2 = model$residuals[lineplotdata$ind == "V2"]
r3 = model$residuals[lineplotdata$ind == "V3"]
r4 = model$residuals[lineplotdata$ind == "V4"]
r5 = model$residuals[lineplotdata$ind == "V5"]
r6 = model$residuals[lineplotdata$ind == "V6"]

# saving into dataframe
residuals1 = data.frame(r1,r2,r3,r4,r5,r6)

```


3.d. Make a lineplot of the realignment parameters with time removed.
```{r}

# lineplot of residuals
lineplot2data <- stack(residuals1, c(r1,r2,r3,r4,r5,r6))
lineplot2data$session = c(1:400, 1:400, 1:400, 1:400, 1:400, 1:400)
lineplot2 <- ggplot(lineplot2data, aes(session, values, group = ind, colour = ind)) + geom_line()

# lineplot parametres with time removed
lineplot2

```
3.e. Make a lineplot including only the first realignment parameter before and after removal.
```{r}

# defining before and after residual removal
After = model$residuals[lineplotdata$ind == "V1"]
Before = parametres[,1]

# constructing a lineplot
lineplot3data = data.frame(Before, After)
lineplot3data <- stack(lineplot3data, c(Before, After))
lineplot3data$session = c(1:400)
lineplot3 <- ggplot(lineplot3data, aes(session, values, group = ind, colour = ind)) +labs(title = "First realignment parameter, before and after removal of the effect of time", x= "Time", y = "Value", colour = "") + geom_line()

# lineplot before and after removal of time effect
lineplot3

```
3.f. Are the realignment parameters (corrected for effects of time) now correlated with the fMRI model?
```{r}

# testing for correlation
corr.test(residuals1, fmrides)

```
No, none of the realignment parametres are correlated with the fMRI model, p=1